Phylogenetic adaptions to urban living in Australian birds

### 2018-04-07 ms.Rmd
flushing out intro

# INTRODUCTION
Urbanization fragments and isolates natural habitat [@marzluff2001restoration; @melles2003urban; @loram2007urban], thus globally threatening biodiversity [@mckinney2006urbanization; @mckinney2008effects]. Despite the documented loss of biodiversity directly or indirectly caused by urbanization [@mckinney1999biotic; @mckinney2006urbanization], some species have successfully adapted to, colonized and subsequently thrived in urban environments [@mckinney2002urbanization; @chace2006urban; @lowry2013behavioural; @mcdonnell2015adaptation].

Within urban ecosystems there are generally urban avoiders, urban adapters, and urban exploiters [@mckinney2002urbanization; @mckinney2006urbanization; @kark2007living; @mcdonnell2015adaptation]. Urban avoiders are sensitive to urban environments [@mcdonnell2015adaptation] and tend to apportion more energy towards reproduction, relative to adapters and exploiters [@croci2008does]. Urban adapters can utilize urban environmental resources but may not reside in them [@mcdonnell2015adaptation] and are generally sedentary, social, high fecundity, with wide distribution [@kark2007living, @croci2008does]. Urban exploiters thrive in urban environments [@mcdonnell2015adaptation] and are generally more social and sedentary than adapters [@kark2007living]. Species’ traits such as social status, behavioral flexibility, migratory status, nesting, diet preferences, habitat preferences, and overall species range have all been used to predict presence in urban environments [@mcclure1989characterizes; @mills1989effects; @jokimaki1998distribution; @lim2004responses; @clergeau2006avifauna; @chace2006urban; @lowry2013behavioural]. Understanding which favored traits are associated with urban adapters and avoiders [@croci2008does] and their habitats can inform management decisions for conservation of particular species and improvements of habitats within urban environments.

@@

Despite the number of studies which have investigated the relationships between species' traits and urban adaptedness, most studies are either spatially limited, such as a single city or region [@blair1996land; @kark2007living; @croci2008does]; or temporally limited to a single breeding season [@blair1996land] or two month survey period [@kark2007living]]. Moreover, the terminology (i.e., urban adapters, avoiders, and exploiters) is used differently among studies [cf., [@kark2007living] and [@croci2008does]], potentially affecting interpretation between scientists and the public [@mcdonnell2015adaptation; @fischer2015categorizing].

>>>>

# INTRODUCTION
Urbanization fragments and isolates natural habitat [@marzluff2001restoration; @melles2003urban; @loram2007urban], drastically altering species' composition and ecosystem functions. Thus, urbanization globally threatens biodiversity through processes such as biotic homogenization [@mckinney2006urbanization; @mckinney2008effects] and is a major threat to conservation [@czech2000economic; @dearborn2010motivations]. Despite the documented loss of biodiversity directly or indirectly caused by urbanization [@mckinney1999biotic; @mckinney2006urbanization; @devictor2007functional], some species have successfully adapted to, colonized and subsequently thrived in urban environments [@mckinney2002urbanization; @chace2006urban; @lowry2013behavioural; @mcdonnell2015adaptation]. For instance, some species which rarely occur within urban areas can be relatively common in surrounding rural habitats [@tratalos2007bird; @evans2011makes; @gatesire2014bird].

In order to best predict species' responses to future urbanization changes, it is essential to understand what biological traits are associated with urban birds [e.g., @mcclure1989characterizes; @bonier2007urban; @moller2009successful;  @mcdonnell2015adaptation]. To date, a wide array of traits have been used to assess presence of birds in urban environments. For instance, degree of sociality [@mcclure1989characterizes; @coleman1994neophobia; @jokimaki1998distribution], nesting substrate (CITATION), diet (CITATIONS), distribution size [@jokimaki2000artificial; @chace2006urban; @croci2008does], migratory status [@friesen1995effects], niche breadth [@kark2007living; @evans2011make], and fecundity [@croci2008does; @moller2009successful; @vaugoyeau2016interspecific], among other traits (see Appendix 1), have all been hypothesized to affect a birds' presence in urban environments. However, a wide range of results have been reported in the literature, often contrasting one another. Brain size was found to be important by [@moller2015brain & @maklakov2011brains] but [@evans2011make & @kark2007living] found that brain size had no effect on species' response to urbanization. Similarly, contrasting results have been reported for annual fecundity [cf., @evans2011make with @moller2009successful & @croci2008does] and generalism or niche breadth [cf., @kark2007living with @evans2011make].

The majority of previous studies place species in discrete groups such as urban avoiders, adapters, or exploiters [e.g., @mckinney2002urbanization; @mckinney2006urbanization; @kark2007living; @mcdonnell2015adaptation] or even birds which do or do not occur within urban environments [e.g., @bonier2007urban; @moller2009successful]. This severely limits the generalization of these studies' results as it assumes species within a group respond to urbanization equally [@evans2011make]. Moreover, the terminology (i.e., urban adapters, avoiders, and exploiters) is used differently among studies [cf., [@kark2007living] and [@croci2008does]], potentially affecting interpretation between scientists and the public [@mcdonnell2015adaptation; @fischer2015categorizing]. In a notable exception, [@evans2011make] advocates for a continuous measure of urbanization in which a species' is assessed on their entire distribution in response to urban environments.

Despite the number of studies which have investigated the relationships between species' traits and urban adaptedness, most studies are either spatially limited, such as a single city or region [@blair1996land; @kark2007living; @croci2008does]; or temporally limited to a single breeding season [@blair1996land] or even two month survey period [@kark2007living]]. Most studies also neglect any phylogenetic constraints in their study design [but see @bonier2007urban; @evans2011make], or their analyses are conducted within a single closely-related group of birds [@maklakov2011brains]. We note that the generally small sample sizes (i.e., number of species) might not require the use such phylogenetically-controlled analyses.

Urban avoiders are sensitive to urban environments [@mcdonnell2015adaptation] and tend to apportion more energy towards reproduction, relative to adapters and exploiters [@croci2008does]. Urban adapters can utilize urban environmental resources but may not reside in them [@mcdonnell2015adaptation] and are generally sedentary, social, high fecundity, with wide distribution [@kark2007living, @croci2008does]. Urban exploiters thrive in urban environments [@mcdonnell2015adaptation] and are generally more social and sedentary than adapters [@kark2007living]. Species’ traits such as social status, behavioral flexibility, migratory status, nesting, diet preferences, habitat preferences, and overall species range have all been used to predict presence in urban environments [@mcclure1989characterizes; @mills1989effects; @jokimaki1998distribution; @lim2004responses; @clergeau2006avifauna; @chace2006urban; @lowry2013behavioural]. Understanding which favored traits are associated with urban adapters and avoiders [@croci2008does] and their habitats can inform management decisions for conservation of particular species and improvements of habitats within urban environments.

@@
